---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mapnhanespa

<!-- badges: start -->
[![R-CMD-check](https://github.com/jhuwit/mapnhanespa/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jhuwit/mapnhanespa/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/jhuwit/mapnhanespa/graph/badge.svg)](https://app.codecov.io/gh/jhuwit/mapnhanespa)
[![CRAN status](https://www.r-pkg.org/badges/version/mapnhanespa)](https://CRAN.R-project.org/package=mapnhanespa)
<!-- badges: end -->

`mapnhanespa` maps physical activity summaries from a study sample onto
population-level quantiles estimated from NHANES accelerometer data.

## Installation

You can install the development version of mapnhanespa from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("jhuwit/mapnhanespa")
```

## Example

Map one row per participant-measure observation with
`map_nhanes_pa_quantiles()`:

```{r example}
library(mapnhanespa)

study_data <- data.frame(
  id = c("P01", "P02", "P03"),
  age = c(25, 62, 84),
  sex = c("Female", "Male", "Female"),
  measure = c("mims", "ssl_steps", "AC"),
  value = c(15000, 7500, 1000000)
)

map_nhanes_pa_quantiles(study_data, id = "id")
```

The returned `age_category_pa_map` column shows the NHANES age category used
to select each participant's CDF. The following example covers all nine
supported summaries:

```{r all-measures}
all_measures <- data.frame(
  id = rep("P01", 9),
  age = 25,
  sex = "Female",
  measure = c("total_activity_counts", "total_log10AC", "total_PAXMTSM",
              "total_log10PAXMTSM", "total_scsslsteps", "total_scrfsteps",
              "total_oaksteps", "total_vssteps", "total_vsrevsteps"),
  value = c(2773035, 3039, 14653, 1043, 8694, 11421, 12055, 9505, 9631)
)

map_nhanes_pa_quantiles(all_measures, id = "id")
```

The complete list of accepted names and the summary used for each CDF is
available from `nhanes_pa_measure_map()`. Other names return `NA` with a warning listing the
unmapped measures. Each `value` should be calculated using the summary named
in the map and supplied on that summary's scale.

```{r measure-map}
nhanes_pa_measure_map()
```

The `measure` column accepts common aliases, including `total_ac` and
`total_activity_counts` for the `AC` CDF:

```{r aliases}
measures <- data.frame(
  id = rep("P01", 4),
  age = 25,
  sex = "Female",
  measure = c("mims", "PAXMTSM", "total_ac", "total_activity_counts"),
  value = c(15000, 15000, 2773035, 2773035)
)

map_nhanes_pa_quantiles(measures, id = "id")

# Select a CDF function directly for a supported measure and stratum.
nhanes_pa_measure_cdf("total_activity_counts", age = 25, sex = "Female")(2773035)
```

By default, quantiles are evaluated against the combined 2011-2012 and
2013-2014 NHANES waves:

```{r combined}
map_nhanes_pa_quantiles(study_data, id = "id")
```

To map against a specific NHANES wave, provide `wave`:

```{r wave}
map_nhanes_pa_quantiles(study_data, id = "id", wave = "2013-2014")
```

You can also map without sex or age stratification:

```{r strata}
map_nhanes_pa_quantiles(study_data, id = "id", sex = NULL)
map_nhanes_pa_quantiles(study_data, id = "id", age = NULL)
```

For a single participant-measure value, use `nhanes_pa_quantile()`:

```{r single}
nhanes_pa_quantile(
  value = 15000,
  age = 25,
  sex = "Female",
  measure = "mims"
)
```

If a study already has age categories, pass the column name through
`age_category`.

## Funding acknowledgement

This work was supported by NIH P30AG021334 and the Johns Hopkins Older Americans Independence Center (OAIC) Pepper Center grant.
